---
title: "Data Analysis"
output:
  workflowr::wflow_html:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

2. Data Analysis
    i) [Patterns of Herbicide Use](analysis.html)
    ii) [Changes in Land Cover](analysis_land.html)
    iii) [Identifying Herbicide Active Ingredients](analysis_ingredient.html)
    
    
#**ii) Changes in Land Cover**
```{r message=FALSE, echo=FALSE}
library(tidyverse)
library(plyr)
library(dplyr)
library(png)
library(grid)
library(readxl)
library(reshape2)
```

All the files created in *Data Preparation* was imported. While the focus on analysis in this page was with land cover changes, the other files would help with joining and normalizing the dataframes.
```{r}
lc97 <- read_excel("../data/big_data/lc97.xlsx")
lc02 <- read_excel("../data/big_data/lc02.xlsx")
lc07 <- read_excel("../data/big_data/lc07.xlsx")
lc12 <- read_excel("../data/big_data/lc12.xlsx")
county <- read_excel("../data/big_data/county.xlsx")
herbicide_use <- read_excel("../data/big_data/herbicide_use.xlsx")
```

Out of the 17 land cover properties, 4 were selected. These 4 land cover were estimated to influence the availability of nectar for the moanrchs:

* **Grassland**
* **Shrubland**
* **Cropland**
* **Hay/Pasture**

***

###a. Calculating Percent of Land cover
Before calculating the changes of land cover of the selected land cover properties, all the data from differnet land cover needed to be organized into one table.

The land cover dataframes were cleaned.
```{r}
lc97_select <- lc97 %>% 
  select(2, 13:16, 20)
lc02_select <- lc02 %>% 
  select(2, 13:16, 20)
lc07_select <- lc07 %>% 
  select(2, 13:16, 20)
lc12_select <- lc12 %>% 
  select(2, 13:16, 20)
```

The land cover properties were then normalized by the *sum*.
```{r}
lc97_select <- lc97_select %>% 
  mutate(grass97=(grassland/sum)*100, shrub97=(shrubland/sum)*100, crop97=(cropland/sum)*100, hay97=(hay_pasture_land/sum)*100) %>% 
  select(1, 7:10)
lc02_select <- lc02_select %>% 
  mutate(grass02=(grassland/sum)*100, shrub02=(shrubland/sum)*100, crop02=(cropland/sum)*100, hay02=(hay_pasture_land/sum)*100) %>% 
  select(1, 7:10)
lc07_select <- lc07_select %>% 
  mutate(grass07=(grassland/sum)*100, shrub07=(shrubland/sum)*100, crop07=(cropland/sum)*100, hay07=(hay_pasture_land/sum)*100) %>% 
  select(1, 7:10)
lc12_select <- lc12_select %>% 
  mutate(grass12=(grassland/sum)*100, shrub12=(shrubland/sum)*100, crop12=(cropland/sum)*100, hay12=(hay_pasture_land/sum)*100) %>% 
  select(1, 7:10)
```

Finally these dataframed were merged into one **lc4**. 
```{r}
lc4 <- merge(x=lc97_select, y=lc02_select, by="fips")
lc4 <- merge(x=lc4, y=lc07_select, by="fips")
lc4 <- merge(x=lc4, y=lc12_select, by="fips")
```

This merged data was then merged again to the *county* table so that all the percent value could be represented in acres. 
```{r}
lc4 <- merge(x=county, y=lc4, by="fips")
```

The perent values were then calculated to show land cover properties in acres.
```{r}
lc4_acre <- lc4 %>%
  mutate(g97_acre=(grass97/100)*area_acre, s97_acre=(shrub97/100)*area_acre, c97_acre=(crop97/100)*100, h97_acre=(hay97/100)*area_acre, g02_acre=(grass02/100)*area_acre, s02_acre=(shrub02/100)*area_acre, c02_acre=(crop02/100)*100, h02_acre=(hay02/100)*area_acre, g07_acre=(grass07/100)*area_acre, s07_acre=(shrub07/100)*area_acre, c07_acre=(crop07/100)*100, h07_acre=(hay07/100)*area_acre, g12_acre=(grass12/100)*area_acre, s12_acre=(shrub12/100)*area_acre, c12_acre=(crop12/100)*100, h12_acre=(hay12/100)*area_acre) %>% 
  select(1,3,21:36)
```

Now that all land cover properties were shown in acres, each land cover were made into its own data frams so that further analysis could take place.
```{r}
grass <- lc4_acre %>% 
  select(1,2,3,7,11,15)
shrub <- lc4_acre %>% 
  select(1,2,4,8,12,16)
crop <- lc4_acre %>% 
  select(1,2,5,9,13,17)
hay <- lc4_acre %>% 
  select(1,2,6,10,14,18)
```

***

###b. Change from 1997 to 2012
Like herbicide use, the change from 1997 to 2012 was done for all the chosen land cover.
```{r}
grass <- grass %>% 
  mutate(grass_chg=(g12_acre-g97_acre))
shrub <- shrub %>% 
  mutate(shrub_chg=(s12_acre-s97_acre))
crop <- crop %>% 
  mutate(crop_chg=(c12_acre-c97_acre))
hay <- hay %>% 
  mutate(hay_chg=(h12_acre-h97_acre))
```

This change was joined with *county* dataframe so that each change could be normalized with the size of the county. 
```{r}
grass1 <- merge(x=county, y=grass, by="fips") 
shrub1 <- merge(x=county, y=shrub, by="fips")
crop1 <- merge(x=county, y=crop, by="fips")
hay1 <- merge(x=county, y=hay, by="fips")
```

The change was then normalized. 
```{r}
grass1 <- grass1 %>% 
  mutate(lg_grassch=(grass_chg/area_acre)*100)
shrub1 <- shrub1 %>% 
  mutate(lg_shrubch=(shrub_chg/area_acre)*100)
crop1 <- crop1 %>% 
  mutate(lg_cropch=(crop_chg/area_acre)*100)
hay1 <- hay1 %>% 
  mutate(lg_haych=(hay_chg/area_acre)*100)
```

***
```{r, echo=FALSE}
herb_acre <- merge(x=herbicide_use, y=county, by="fips")
```
```{r, echo=FALSE}
herb_acre <- herb_acre %>% 
  mutate(herb97_acre=herbicide_97/area_acre, herb02_acre=herbicide_02/area_acre, herb07_acre=herbicide_07/area_acre, herb12_acre=herbicide_12/area_acre) %>% 
  select(1,3,4,11:15)
```
```{r, echo=FALSE}
herb_acre1 <- herb_acre %>% 
  mutate(herb_change=((herb12_acre-herb97_acre)/herb97_acre)*100)
```
```{r, echo=FALSE}
herb_state <- ddply(herbicide_use, "State", numcolwise(sum), na.rm=TRUE)
```
```{r, echo=FALSE}
state_acre = ddply(herb_acre, "State", numcolwise(sum))
state_acre <- state_acre %>% 
  select(1,2)
```
```{r, echo=FALSE}
herb_state = merge(x=herb_state, y=state_acre, by="State")
```
```{r, echo=FALSE}
herb_state <- herb_state %>% 
  mutate(herb97_st=herbicide_97/area_acre, herb02_st=herbicide_02/area_acre, herb07_st=herbicide_07/area_acre, herb12_st=herbicide_12/area_acre)
```


###c. Trend in land cover changes
The land cover changes were also prepared to be analized in a state level. The dataframe previously created was joined. 
```{r}
state_join <- herbicide_use %>% 
  select(2,3)
```

The states were joined with the land cover dataframe. These were joined by the key **fips**. 
```{r}
grass_state <- merge(x=state_join, y=grass, by="fips") 
shrub_state <- merge(x=state_join, y=shrub, by="fips")
crop_state <- merge(x=state_join, y=crop, by="fips")
hay_state <- merge(x=state_join, y=hay, by="fips")
```

The land cover properties were aggregate by summing these value in a state level. 
```{r}
grass_state <- ddply(grass_state, "State", numcolwise(sum))
shrub_state <- ddply(shrub_state, "State", numcolwise(sum))
crop_state <- ddply(crop_state, "State", numcolwise(sum))
hay_state <- ddply(hay_state, "State", numcolwise(sum))
```

The values were normalized with the size of the state.
```{r}
grass_state <- merge(x=state_acre, y=grass_state, by="State")
shrub_state <- merge(x=state_acre, y=shrub_state, by="State")
crop_state <- merge(x=state_acre, y=crop_state, by="State")
hay_state <- merge(x=state_acre, y=hay_state, by="State")
```

A percent change was calculated:
```{r}
grass_state <- grass_state %>%  
  mutate(g97_per=(g97_acre/area_acre)*100, g02_per=(g02_acre/area_acre)*100, g07_per=(g07_acre/area_acre)*100, g12_per=(g12_acre/area_acre)*100)
shrub_state <- shrub_state %>%  
  mutate(s97_per=(s97_acre/area_acre)*100, s02_per=(s02_acre/area_acre)*100, s07_per=(s07_acre/area_acre)*100, s12_per=(s12_acre/area_acre)*100)
crop_state <- crop_state %>%  
  mutate(c97_per=(c97_acre/area_acre)*100, c02_per=(c02_acre/area_acre)*100, c07_per=(c07_acre/area_acre)*100, c12_per=(c12_acre/area_acre)*100)
hay_state <- hay_state %>%  
  mutate(h97_per=(h97_acre/area_acre)*100, h02_per=(h02_acre/area_acre)*100, h07_per=(h07_acre/area_acre)*100, h12_per=(h12_acre/area_acre)*100)
```

Calculating percent change from base year 1997. 
```{r}
grass_state1 <- grass_state %>%
  mutate(g97_base=0, g02_base=g02_per-g97_per, g07_base=g07_per-g97_per, g12_base=g12_per-g97_per)
shrub_state1 <- shrub_state %>%
  mutate(s97_base=0, s02_base=s02_per-s97_per, s07_base=s07_per-s97_per, s12_base=s12_per-s97_per)
crop_state1 <- crop_state %>%
  mutate(c97_base=0, c02_base=c02_per-c97_per, c07_base=c07_per-c97_per, c12_base=c12_per-c97_per)
hay_state1 <- hay_state %>%
  mutate(h97_base=0, h02_base=h02_per-h97_per, h07_base=h07_per-h97_per, h12_base=h12_per-h97_per)
```

The tables were refined to show only the relevant columns. 
```{r}
grass_state1 <- grass_state1 %>% 
  select(-2:-11)
shrub_state1 <- shrub_state1 %>% 
  select(-2:-11)
crop_state1 <- crop_state1 %>% 
  select(-2:-11)
hay_state1 <- hay_state1 %>% 
  select(-2:-11)
```
***

###d. Land Use Changes (Table)
Preparing table for land use.

Changing column names for join. 
```{r}
names(lc97)[3] <-  "water97"
names(lc97)[4] <-  "developed97"
names(lc97)[5] <- "national_forest97"
names(lc97)[6] <- "public_land97"
names(lc97)[7] <- "private97"
names(lc97)[8] <- "mining97"
names(lc97)[9] <- "barren_97"
names(lc97)[10] <- "deci_forest97"
names(lc97)[11] <- "ever_forest97"
names(lc97)[12] <- "mix_forest97"
names(lc97)[13] <- "grass97"
names(lc97)[14] <- "shrub97"
names(lc97)[15] <- "crop97"
names(lc97)[16] <- "hay97"
names(lc97)[17] <- "herbaceous97"
names(lc97)[18] <- "wetland97"
names(lc97)[19] <- "ice97"
names(lc97)[20] <- "sum97"

names(lc12)[3] <-  "water12"
names(lc12)[4] <-  "developed12"
names(lc12)[5] <- "national_forest12"
names(lc12)[6] <- "public_land12"
names(lc12)[7] <- "private12"
names(lc12)[8] <- "mining12"
names(lc12)[9] <- "barren_12"
names(lc12)[10] <- "deci_forest12"
names(lc12)[11] <- "ever_forest12"
names(lc12)[12] <- "mix_forest12"
names(lc12)[13] <- "grass12"
names(lc12)[14] <- "shrub12"
names(lc12)[15] <- "crop12"
names(lc12)[16] <- "hay12"
names(lc12)[17] <- "herbaceous12"
names(lc12)[18] <- "wetland12"
names(lc12)[19] <- "ice12"
names(lc12)[20] <- "sum12"
```

Joining the two lc data, and removing extra data. 
```{r}
lc_table=left_join(lc97, lc12, by="fips")
lc_table <- lc_table %>% 
  select(-21)
```

Joining with county and herbicide_use
```{r}
lc_table=left_join(lc_table, county, by="fips")
lc_table=left_join(lc_table, herbicide_use, by="fips")

lc_table <- lc_table %>% 
  select(-1,-2,-39,-40,-42,-44:-48)
```

Aggregating
```{r}
lc_table_st <- ddply(lc_table, "State", numcolwise(sum))
```

Calculating percent.
```{r}
lc_table_st97 <- lc_table_st %>% 
  mutate(Water97=(water97/sum97)*100, Developed97=(developed97/sum97)*100, National_forest97=(national_forest97/sum97)*100, Public_land97=(public_land97/sum97)*100, Private97=(private97/sum97)*100, Mining97=(mining97/sum97)*100, Barren97=(barren_97/sum97)*100, Deci_forest97=(deci_forest97/sum97)*100, Ever_forest=(ever_forest97/sum97)*100, Mix_forest=(mix_forest97/sum97)*100, Grass97=(grass97/sum97)*100, Shrub97=(shrub97/sum97)*100, Crop97=(crop97/sum97)*100, Hay97=(hay97/sum97)*100, Herbaceous97=(herbaceous97/sum97)*100, Wetland97=(wetland97/sum97)*100, Ice97=(ice97/sum97)*100) 
```
```{r}
lc_table_st12 <- lc_table_st %>%
  mutate(Water12=(water12/sum12)*100, Developed12=(developed12/sum12)*100, National_forest12=(national_forest12/sum12)*100, Public_land12=(public_land12/sum12)*100, Private12=(private12/sum12)*100, Mining12=(mining12/sum12)*100, Barren12=(barren_12/sum12)*100, Deci_forest12=(deci_forest12/sum12)*100, Ever_forest=(ever_forest12/sum12)*100, Mix_forest=(mix_forest12/sum12)*100, Grass12=(grass12/sum12)*100, Shrub12=(shrub12/sum12)*100, Crop12=(crop12/sum12)*100, Hay12=(hay12/sum12)*100, Herbaceous12=(herbaceous12/sum12)*100, Wetland12=(wetland12/sum12)*100, Ice12=(ice12/sum12)*100)
```

*** 

