---
title: "Data Analysis"
output:
  workflowr::wflow_html:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

2. Data Analysis
    i) [Patterns of Herbicide Use](analysis.html)
    ii) [Changes in Land Cover](analysis_land.html)
    iii) [Identifying Herbicide Active Ingredients](analysis_ingredient.html)
    
#Incomplete
    
#**iii) Identifying Herbicide Active Ingredients**
```{r message=FALSE, echo=FALSE}
library(tidyverse)
library(plyr)
library(dplyr)
library(png)
library(grid)
library(readxl)
library(reshape2)
```

Importing the excel file to R.
```{r}
join_herb2 <- read_excel("../data/big_data/join_herb2.xlsx")
```

Aggregating for **Cropland** and **Hay/Pasture Land**.
```{r}
ing_sum <- join_herb2 %>% 
  select(-Units, -compound_std, -category) %>% 
  mutate(crop_sum=select(., `Corn`:`Alfalfa`, `Other_crops`) %>% 
           apply(1, sum, na.rm=TRUE))
```
###a. Overall Site: Crop vs Hay trends
Just selecting crop vs hay
```{r}
ing_crop_hay <- ing_sum %>% 
  select(-1:-4, -6:-12, -15) %>% 
  ddply(., "Year", numcolwise(sum), na.rm=TRUE)
```
Finding the ratio (mutate ratio)
```{r}
ing_crop_hay <- ing_crop_hay %>% 
  mutate(ratio=Pasture_and_hay/crop_sum)
```

###b. Active Ingredient percent change from 1997 to 2012
Just selecting the ingredients. First one for 1997
```{r}
ing97_only <- ing_sum %>% 
  select(-1:-3,-6:-13,-15) %>% 
  filter(Year == 1997) %>% 
  ddply(., "Compound", numcolwise(sum), na.rm=TRUE)

names(ing97_only)[3] <- "hay97"  
names(ing97_only)[4] <- "crop97"
```

```{r}
ing12_only <- ing_sum %>% 
  select(-1:-3,-6:-13,-15) %>% 
  filter(Year == 2012) %>% 
  ddply(., "Compound", numcolwise(sum), na.rm=TRUE)

names(ing12_only)[3] <- "hay12"  
names(ing12_only)[4] <- "crop12"
```

Combining the two dataframes
```{r}
ing_97_12 = merge(x=ing97_only, y=ing12_only, by="Compound")
```

Calculations
``{r}
ing_97_12 <- ing_97_12 %>% 
  select(-2,-5) %>% 
  mutate(sum97=hay97+crop97, sum12=hay12+crop12, ing_per=((sum12-sum97)/sum97)*100)
```




