---
title: "Data Analysis"
output:
  workflowr::wflow_html:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

2. Data Analysis
    i) [Patterns of Herbicide Use](analysis.html)
    ii) [Changes in Land Cover](analysis_land.html)
    iii) [Identifying Herbicide Active Ingredients](analysis_ingredient.html)
    
##**This section has been omitted**

#**iii) Identifying Herbicide Active Ingredients**
```{r message=FALSE, echo=FALSE}
library(tidyverse)
library(plyr)
library(dplyr)
library(png)
library(grid)
library(readxl)
library(reshape2)
```

The excel file prepared in the *Data Preparation* was imported again. 
```{r}
join_herb2 <- read_excel("../data/big_data/join_herb2.xlsx")
```

Aggregating for **Cropland** and **Hay/Pasture Land**.
```{r}
ing_sum <- join_herb2 %>% 
  select(-Units, -compound_std, -category) %>% 
  mutate(crop_sum=select(., `Corn`:`Alfalfa`, `Other_crops`) %>% 
           apply(1, sum, na.rm=TRUE))
```

###a. Active Ingredient percent change from 1997 to 2012

Just selecting the ingredients. First one for 1997
```{r}
ing97_only <- ing_sum %>% 
  select(-1:-3,-6:-13,-15) %>% 
  filter(Year == 1997) %>% 
  ddply(., "Compound", numcolwise(sum), na.rm=TRUE)

names(ing97_only)[3] <- "hay97"  
names(ing97_only)[4] <- "crop97"
```

```{r}
ing12_only <- ing_sum %>% 
  select(-1:-3,-6:-13,-15) %>% 
  filter(Year == 2012) %>% 
  ddply(., "Compound", numcolwise(sum), na.rm=TRUE)

names(ing12_only)[3] <- "hay12"  
names(ing12_only)[4] <- "crop12"
```

Combining the two dataframes
```{r}
ing_97_12 = merge(x=ing97_only, y=ing12_only, by="Compound")
```

Calculations
```{r}
ing_97_12 <- ing_97_12 %>% 
  select(-2,-5) %>% 
  mutate(sum97=hay97+crop97, sum12=hay12+crop12, ing_per=((sum12-sum97)/sum97)*100, crop_per=((crop12-crop97)/crop97)*100, hay_per=((hay12-hay97)/hay97)*100, crop_diff=crop12-crop97, hay_diff=hay12-hay97)
```

***

###b. Active Ingredient Map
The data downloaded from the *Data Preparation* section was imported into R.
```{r}
pest_97 <- read_excel("../data/big_data/pest_97.xlsx")
pest_12 <- read_excel("../data/big_data/pest_12.xlsx")
```

Also the *county_area*, which was prepared previously, was imported. This would help in normalizing the values.
```{r}
county <- read_excel("../data/big_data/county_area.xlsx")
```

Filterout only relevant columns: 
```{r}
pest_97 <- pest_97 %>%
  select(1, 5, 6)
names(pest_97)[3] <- "lowkg_97"
pest_12 <- pest_12 %>%
  select(1, 5, 6)
names(pest_12)[3] <- "lowkg_12"
county <- county %>% 
  select(6,7)
names(county)[1] <- "fips"
```

Cannot be converted to excel, maybe too many rows. So trying to aggregate them now. 
```{r}
pest97_glyphosate <- pest_97 %>% 
  filter(COMPOUND=="GLYPHOSATE") %>% 
  ddply(.,"fips", numcolwise(sum), na.rm=TRUE)
names(pest97_glyphosate)[2] <- "glyphosate97"

pest12_glyphosate <- pest_12 %>% 
  filter(COMPOUND=="GLYPHOSATE") %>% 
  ddply(.,"fips", numcolwise(sum), na.rm=TRUE)
names(pest12_glyphosate)[2] <- "glyphosate12"
```

Joining the two tables
```{r}
glyphosate=merge(x=pest97_glyphosate, y=pest12_glyphosate, by="fips")
glyphosate=merge(x=glyphosate, y=county, by="fips")
```
 
 Normalizing
```{r}
glyphosate <- glyphosate %>% 
  mutate(change=((glyphosate12-glyphosate97)/area_acre)*100)
```

Similarly, same step was done for 24_D, Atrazine, Dicamba, and Paraquat. 
```{r, echo=FALSE}
pest97_24_D <- pest_97 %>% 
  filter(COMPOUND=="2,4-D") %>% 
  ddply(.,"fips", numcolwise(sum), na.rm=TRUE)
names(pest97_24_D)[2] <- "24_D97"

pest12_24_D <- pest_12 %>% 
  filter(COMPOUND=="2,4-D") %>% 
  ddply(.,"fips", numcolwise(sum), na.rm=TRUE)
names(pest12_24_D)[2] <- "24_D12"

i24_D=left_join(pest97_24_D, pest12_24_D, by="fips")
i24_D=left_join(i24_D, county, by="fips")
```

```{r, echo=FALSE}
pest97atrazine <- pest_97 %>% 
  filter(COMPOUND=="ATRAZINE") %>% 
  ddply(.,"fips", numcolwise(sum), na.rm=TRUE)
names(pest97atrazine)[2] <- "24_D97"

pest12atrazine <- pest_12 %>% 
  filter(COMPOUND=="ATRAZINE") %>% 
  ddply(.,"fips", numcolwise(sum), na.rm=TRUE)
names(pest12atrazine)[2] <- "24_D12"

atrazine=left_join(pest97atrazine, pest12atrazine, by="fips")
atrazine=left_join(atrazine, county, by="fips")
```

```{r, echo=FALSE}
pest97dicamba <- pest_97 %>% 
  filter(COMPOUND=="DICAMBA") %>% 
  ddply(.,"fips", numcolwise(sum), na.rm=TRUE)
names(pest97dicamba)[2] <- "24_D97"

pest12dicamba <- pest_12 %>% 
  filter(COMPOUND=="DICAMBA") %>% 
  ddply(.,"fips", numcolwise(sum), na.rm=TRUE)
names(pest12dicamba)[2] <- "24_D12"

dicamba=left_join(pest97dicamba, pest12dicamba, by="fips")
dicamba=left_join(dicamba, county, by="fips")
```

```{r, echo=FALSE}
pest97paraquat <- pest_97 %>% 
  filter(COMPOUND=="PARAQUAT") %>% 
  ddply(.,"fips", numcolwise(sum), na.rm=TRUE)
names(pest97paraquat)[2] <- "24_D97"

pest12paraquat <- pest_12 %>% 
  filter(COMPOUND=="PARAQUAT") %>% 
  ddply(.,"fips", numcolwise(sum), na.rm=TRUE)
names(pest12paraquat)[2] <- "24_D12"

paraquat=left_join(pest97paraquat, pest12paraquat, by="fips")
paraquat=left_join(paraquat, county, by="fips")
```


***


