---
title: "Data Analysis"
output:
  workflowr::wflow_html:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


2. Data Analysis
    i) [Patterns of Herbicide Use](analysis.html)
    ii) [Changes in Land Cover](analysis_land.html)
    iii) [Identifying Herbicide Active Ingredients](analysis_ingredient.html)
    
    
#**i) Patterns of Herbicide Use**

The majority of the analysis done in this project deals with the three dataset that were cleaned and reorganized in the **Data Preparation** section. Is is important to have the exported excel files of these 3 files that were prepared before.
```{r message=FALSE, echo=FALSE}
library(tidyverse)
library(plyr)
library(dplyr)
library(png)
library(grid)
library(readxl)
library(reshape2)
```

The following excel files were imported. Only these two excel files will be needed to analyse herbicide used per acres. The two excel files used for analysis are: **county.xlsx** and **herbicide_use.xlsx**.  
```{r}
county <- read_excel("../data/big_data/county.xlsx")
herbicide_use <- read_excel("../data/big_data/herbicide_use.xlsx")
```

***

###a. Normalizing Herbicide Use
The two excel files: *herbicide_use* and *counties* were merged. The **area_acres** in the *counties* dataframe will be used to normalize the herbicide used in the county.  
```{r}
herb_acre <- merge(x=herbicide_use, y=county, by="fips")
```

Herbicide applied in acres was divided by the area of the county. The result (herbicide use per acres) was created in a new column for all the years. 
```{r}
herb_acre <- herb_acre %>% 
  mutate(herb97_acre=herbicide_97/area_acre, herb02_acre=herbicide_02/area_acre, herb07_acre=herbicide_07/area_acre, herb12_acre=herbicide_12/area_acre) %>% 
  select(1,3,4,11:15)
```

***

###b. Herbicide use change from 1997 to 2012
First, the change in herbicide use per acre from 1997 to 2012 was calculated. A new column **herb_change** was added which was the percent change of herbicide per acres. 
```{r}
herb_acre1 <- herb_acre %>% 
  mutate(herb_change=(herb12_acre-herb97_acre)*100)
```

***

###c. State level trend in Herbicide Use
Second, the state level trends in herbicide use was calculated. The analysis before was done in a county level. This step deals with a state level analysis. The herbicide use for all the counties of each state was summed.
```{r}
herb_state <- ddply(herbicide_use, "State", numcolwise(sum), na.rm=TRUE)
```

The size of the state in acres was calculated. This would later be used to normalize herbicide use on a state level.  
```{r}
state_acre = ddply(herb_acre, "State", numcolwise(sum))
state_acre <- state_acre %>% 
  select(1,2)
```

This was then merged to the original **herb_state** dataframe.  
```{r}
herb_state = merge(x=herb_state, y=state_acre, by="State")
```

The summed herbicide use was normalized using the area of the state. A new column **herb_st** was created to display the results of this normalization. 
```{r}
herb_state <- herb_state %>% 
  mutate(herb97_st=herbicide_97/area_acre, herb02_st=herbicide_02/area_acre, herb07_st=herbicide_07/area_acre, herb12_st=herbicide_12/area_acre)
```
This final dataset was converted into excel as **herbicide_state.xlsx** and saved into the repository.   
***

