---
title: "Data Preparation"
output:
  workflowr::wflow_html:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Data Preparation
    i) [Herbicide Use](dprep.html)
    ii) [Land Cover](dprep_land.html)
    iii) [County Area / FIPS Codes](dprep_county.html)
    iv) [Herbicide Active Ingredients](dprep_ingredient.html)
    
    
#**iii) County Area / FIPS Code**

```{r message=FALSE, echo=FALSE}
library(tidyverse)
library(plyr)
library(dplyr)
library(png)
library(grid)
library(readxl)
```

To help normalize the *herbicide use* and the *land cover* data with the area of the county, another data with county area (acres) was also prepared. This data had the *FIPS* code, which would later be used to join data and create maps in ArcGIS.

Calculating the county area in acres was done through ArcGIS. First the shapefile with all relevant states was opened. Calculations, such as finding out the acres of a county, can only be done in an *projected coordinate system* which was set to: **USA_Contiguous_Albers_Equal_Area_Conic**. 

Another column in the *attribute table* of the shapefile was added using the **Add Field** tool under **Table Options**. The new columns was named *area_acres* and using another tool: **Calculate Geometry**, the area of each county was calculate . The resulting table was then saved into a *geodatabase* and then exported to an excel file using the **Table to Excel (Conversion)** tool. 

Cleaning this data was then done through R.
The newly created table was imported.
```{r}
county <- read_excel("../data/big_data/county_area.xlsx")
```
Only the relevant columns: *NAME*, *FIPS*, and *area_acre* were selected. 
```{r}
county <- county %>% 
  select(NAME, FIPS, area_acre)
```
*FIPS* was then renamed to *fips* so it would match the key's of other table. 
```{r}
names(county)[2] <- "fips"
```
Finally this *county* data was exported to excel to be used in analysis. The output excel was named **county.xlsx**.    

***
