---
title: "Data Preparation"
output:
  workflowr::wflow_html:
    toc: true
    toc_float:
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Data Preparation
    i) [Herbicide Use](dprep.html)
    ii) [Land Cover](dprep_land.html)
    iii) [County Area / FIPS Codes](dprep_county.html)
    iv) [Herbicide Active Ingredients](dprep_ingredient.html)
    

#**i) Herbicide Use**

###a. Downloading Data
The Herbicide use data was collected from the United States Department of Agriculture (USDA). All of their downloadable data can be filtered from their '[Statistics Service](https://quickstats.nass.usda.gov/)' page. 


The data was selected from the following criteria:

* **Program:** CENSUS
* **Sector:** ENVIRONMENTAL
* **Group:** FARMS & LAND & ASSETS
* **Commodity:** TREATED
* **Data Item:** AG LAND - TREATED, MEASURED IN ACRE
* **Domain:** CHEMICAL, HERBICIDE
* **Geographic Level:** COUNTY

For the **State:** the 7 states within the study sites - *Texas(TX)*, *Oklahoma(OK)*, *Arkansas(AK)*, *Louisiana(LA)*, *Missouri(MO)*, *Kansas(KS)*, and *Nebraska(NE)* were selected.

For the **Year:** all the years available for download (*1997*, *2002*, *2007*, *2012*) were selected and downloaded individually. These files were saved as excel files and were renamed as: "**herb_97**", "**herb_02**", "**herb_07**", and "**herb_12**" respectively into the repository.  

***

###b. Cleaning and Organizing Data
All of the cleaning and organizing the Herbicide data was mostly done through `R`, but there was an important step was done through `Excel`. 

Each data table needs a **fips** column, which is the counties fips code. Fips codes are unique for each county and this column will act as a key when numerous data sets are joined together for analysis. A fips column was created in excel for all the years. The *fips* column replaced the *Region* column and they were the cumulation of *State ANS* and *County ANS*. After doing this, these excel files were ready to be imported to R. 

This is the library used for data manupulation. This is the same library used for all sections that use R.  
```{r message=FALSE}
library(tidyverse)
library(plyr)
library(dplyr)
library(png)
library(grid)
library(readxl)
```

The downloaded excel files were uploaded into R:
```{r load_data}
herb_97 <- read_excel("../data/big_data/herb_97.xlsx")
herb_02 <- read_excel("../data/big_data/herb_02.xlsx")
herb_07 <- read_excel("../data/big_data/herb_07.xlsx")
herb_12 <- read_excel("../data/big_data/herb_12.xlsx")
```

The data contained a lot of columns that were unnecessary for analysis. Apart from the *Year*, *State*, *County*, and *Value*, all other columns were removed:
```{r}
herb_97[,c(1,3:5,7:9,11,12,14:19,21)] <- NULL
herb_02[,c(1,3:5,7:9,11,12,14:19,21)] <- NULL
herb_07[,c(1,3:5,7:9,11,12,14:19,21)] <- NULL
herb_12[,c(1,3:5,7:9,11,12,14:19,21)] <- NULL
```

The column structure type was corrected. (*Value* fields was not numeric) 
```{r warning=FALSE}
herb_97$Value=as.numeric(herb_97$Value)
herb_02$Value=as.numeric(herb_02$Value)
herb_07$Value=as.numeric(herb_07$Value)
herb_12$Value=as.numeric(herb_12$Value)
```

The columns names for *Value* were  changed so that it would be distinguishable once these data were combined into one table. 
```{r}
names(herb_97)[5] <- "herbicide_97"
names(herb_02)[5] <- "herbicide_02"
names(herb_07)[5] <- "herbicide_07"
names(herb_12)[5] <- "herbicide_12"
```

Since the column names were now distinguishable by year, the *Years* and *State* in all the previous data frames were deleted so when these data frames were merged.
```{r}
herb97_join <- herb_97 %>% 
  select(-Year)
herb02_join <- herb_02 %>% 
  select(fips, herbicide_02)
herb07_join <- herb_07 %>%
  select(fips, herbicide_07)
herb12_join <- herb_12 %>% 
  select(fips, herbicide_12)
```

These dataframes were then merged.
```{r}
herbicide_use <- merge(x=herb97_join, y=herb02_join, by="fips")
herbicide_use <- merge(x=herbicide_use, y=herb07_join, by="fips")
herbicide_use <- merge(x=herbicide_use, y=herb12_join, by="fips")
```

Finally this data was exported as another excel file. This was down with the code `write.xlsx(dataframe, "herbicide_use.xlsx", sheetname = "Sheet1")`. This dataset was named: **herbicide_use.xlsx**, and had the following columns: 

fips | State | County | herbicide_97 | herbicide_02 | herbicide_07 | herbicide_12
-----|-------|--------|--------------|--------------|--------------|-------------

***
