---
title: "Data Preparation"
output:
  workflowr::wflow_html:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Data Preparation
    i) [Herbicide Use](dprep.html)
    ii) [Land Cover](dprep_land.html)
    iii) [County Area / FIPS Codes](dprep_county.html)
    iv) [Herbicide Active Ingredients](dprep_ingredient.html)

    
#Lists of Herbicide Active Ingredient
```{r message=FALSE, echo=FALSE}
library(tidyverse)
library(plyr)
library(dplyr)
library(png)
library(grid)
library(readxl)
```

The dataset for the list of herbicide active ingredient was done in two steps. First, a dataset containing various pesticide use for different crops was downloaded. This data is made available by the USGS's **National Water-Quality Assessment (NAWQA) Project**. In their '[Pesticide National Synthesis Project](https://water.usgs.gov/nawqa/pnsp/usage/maps/county-level/)' page, the *"Low Estimate Agricultural Pesticide Use by Corn Group 1992 to 2012"* was downloaded. This dataset contains all the pesticide used in various crops from the years 1992 to 2012 at a county level. Second, in order to filter out the herbicide from the list of pesticides, another dataset containing the list of herbicides compounds was downloaded. 

These dataset were converted into excel and imported into **R**.
```{r warning=FALSE}
active_ing <- read_excel("../data/big_data/pesticideuse_low.xlsx")
usgs_herbicide <- read_excel("../data/big_data/usgs_herbicides.xlsx")
```

The NAWQA's dataset on pesticuse is extensive, so the first step was filtering down this dataset into the states and time frame used for this project's analysis.
```{r}
all_ingredient <- active_ing %>% 
  filter(State %in% c("Arkansas","Kansas","Louisiana","Missouri","Nebraska","Oklahoma","Texas"), Year %in% c(1997, 2002, 2007, 2012))
```

A function to remove extra characters in *compound* name was created. This would help join compounds to the list of herbicides. 
```{r}
cmpd_std <- function(x){toupper(gsub(",|-| ","",x))}
```

The function was applied. 
```{r}
all_ingredient <- all_ingredient %>%
  mutate(compound_std=cmpd_std(Compound))
```

Similarly, only the relevant columns from the *herbicide list* was filtered.   
```{r}
usgs_herb <- usgs_herbicide %>% 
  select(1,2)
names(usgs_herb)[1] <- "compound_std"
```

The *herbicide list* was then joined with the more extensive pesticide data set, so only herbicide data's would show.
```{r}
join_herb <- all_ingredient %>%
  left_join(.,usgs_herb, by="compound_std")
```
All the non-herbicide ingredients were removed. The resulting dataset was converted into excel for further analysis.
```{r}
join_herb2 <- join_herb[-which(is.na(join_herb$category)),]
```

The excel output was named **join_herb2.xlsx** and contains values of broadleaf herbicide compounds used in different crops.  
***

#Herbicide Active Ingredient Map
To figure out the list of herbicide active ingredient used at a county level, another dataset from the same [Pesticide National Synthesis Project](https://water.usgs.gov/nawqa/pnsp/usage/maps/county-level/)'s page was downloaded. The **1997** and the **2012** 'County-level pesticide use estimates' was downloaded. 

After downloading, a *fips* column was created in `Excel`. The fips column was calculated using the **concatenate** formula which was used to join the values of *STATE_FIPS_CODE* and *COUNTY_FIPS_CODE*. The newly created **fips** column would help map the data in ArcGIS.

***
